Efficiently summarising relationships in large samples: a general duality between statistics of genealogies and genomes

### 2019-09-01 treestats_paper.tex
Draft figure showing algorithm state.

%%%%%%%%%%%%%%%%%%%%%
\subsection*{Algorithms}
write-up of algorithm for maintaining branch statistic across trees

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%
\subsection*{Algorithms}
In this section we provide describe and analyse the algorithm used
to compute branch length statistics. This is a generalation
of the algorithm used to maintain the number of samples from a
given set that derive from each node in a tree sequence.
\cite[Algorithm L]{kelleher2016efficient}.

\jk{Plan for section}
\begin{itemize}
\item First paragraph is a general outline of the strategy.
\item Explanation of notation and the index vectors (we should probably define
these again, in the interest of completeness. I'm not sure the definition
in the msprime paper really worked)
\item Algorithm listing.
\item Next para is an explanation of the algorithm in words, referring
to the figure for what we mean by adding and removing edges.
\item Next para is an analysis of the algorithm. We note here that
we propagate stuff all the way up the tree close to optimally, because of the
ordering of edge insertions and removals. For example, we always disconnect
a subtree from the main tree first, meaning that the only time we go all the
way to the root is the first disconnection.
Discuss the figure perf figure.
\end{itemize}

@@

\begin{figure}

% Run the script illustrate_algorithm.py to generate these pdfs and
% the latex tables below.
\begin{tabular}{cccc}
\includegraphics[width=3cm]{figures/tree_0_in_6.pdf} &
\includegraphics[width=3cm]{figures/tree_0_out_0.pdf} &
\includegraphics[width=3cm]{figures/tree_0_out_1.pdf} &
\includegraphics[width=3cm]{figures/tree_1_in_0.pdf}
\\

\begin{tabular}{c|cccc}
& $\pi$ & $\beta$ & $x$ & $F$\\
0 & 7 & 2.0 & 1.0 & 0.2\\
1 & 6 & 1.0 & 1.0 & 0.2\\
2 & 6 & 1.0 & 1.0 & 0.2\\
3 & 6 & 1.0 & 1.0 & 0.2\\
4 & 8 & 3.0 & 1.0 & 0.2\\
5 & -1 & 0.0 & 0.0 & 0.0\\
6 & 7 & 1.0 & 3.0 & 0.3\\
7 & 8 & 1.0 & 4.0 & 0.2\\
8 & -1 & 0.0 & 5.0 & 0.0\\
\end{tabular}
&
\begin{tabular}{c|cccc}
& $\pi$ & $\beta$ & $x$ & $F$\\
0 & -1 & 0.0 & 1.0 & 0.2\\
1 & 6 & 1.0 & 1.0 & 0.2\\
2 & 6 & 1.0 & 1.0 & 0.2\\
3 & 6 & 1.0 & 1.0 & 0.2\\
4 & 8 & 3.0 & 1.0 & 0.2\\
5 & -1 & 0.0 & 0.0 & 0.0\\
6 & 7 & 1.0 & 3.0 & 0.3\\
7 & 8 & 1.0 & 3.0 & 0.3\\
8 & -1 & 0.0 & 4.0 & 0.2\\
\end{tabular}
&
\begin{tabular}{c|cccc}
& $\pi$ & $\beta$ & $x$ & $F$\\
0 & -1 & 0.0 & 1.0 & 0.2\\
1 & 6 & 1.0 & 1.0 & 0.2\\
2 & 6 & 1.0 & 1.0 & 0.2\\
3 & -1 & 0.0 & 1.0 & 0.2\\
4 & 8 & 3.0 & 1.0 & 0.2\\
5 & -1 & 0.0 & 0.0 & 0.0\\
6 & 7 & 1.0 & 2.0 & 0.3\\
7 & 8 & 1.0 & 2.0 & 0.3\\
8 & -1 & 0.0 & 3.0 & 0.3\\
\end{tabular}
&
\begin{tabular}{c|cccc}
& $\pi$ & $\beta$ & $x$ & $F$\\
0 & 5 & 1.0 & 1.0 & 0.2\\
1 & 6 & 1.0 & 1.0 & 0.2\\
2 & 6 & 1.0 & 1.0 & 0.2\\
3 & -1 & 0.0 & 1.0 & 0.2\\
4 & 8 & 3.0 & 1.0 & 0.2\\
5 & -1 & 0.0 & 1.0 & 0.2\\
6 & 7 & 1.0 & 2.0 & 0.3\\
7 & 8 & 1.0 & 2.0 & 0.3\\
8 & -1 & 0.0 & 3.0 & 0.3\\
\end{tabular}
\end{tabular}

\caption{Example of the algorithm state during the transition between two
trees. \jk{If we keep this then we talk about what we see in the different
examples. We want to give some insight into how the algorithm works by showing
the intermediate states.}}
\end{figure}
